### 2014-04-01 ms_probaweb.tex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{color,lineno,setspace,multirow}
\usepackage[sc]{mathpazo}
\usepackage{graphicx}
\usepackage[top=2.4cm,left=2.4cm,top=2.4cm,bottom=2.4cm,includefoot]{geometry}
\usepackage[style=bes, backend=bibtex]{biblatex}
\usepackage{rotating}

\addbibresource{tim.bib}

\begin{document}

\linenumbers 
\modulolinenumbers[1]

\textbf{Title:} A quantitative framework for network biogeography\\
% Alternative: The distribution of ecological interactions 
% Alternative: The biogeography of ecological interactions
% Alternative: Community (macro)ecology: integrating species interactions and species distributions

\textbf{Authors:} Dominique Gravel$^{1,2,*}$, Timoth\'ee Poisot$^{1,2,3}$, N\'eo
Martinez$^{4}$, Jennifer A. Dunne$^{5}$, Jason Tylianakis$^{3}$, Nicolas
Mouquet$^{6}$, Daniel B. Stouffer$^{3}$ \\

1: Canada Research Chair in Biogeography and community ecology. D\'epartement de
biologie, chimie et g\'eographie, Universit\'e du Qu\'ebec \`a Rimouski, 300
All\'ee des Ursulines, Qu\'ebec, Canada. G5L 3A1.\\

2: Qu\'ebec Centre for Biodiversity Sciences\\

3: \\

4: \\

5: \\

6:\\

\textbf{Keywords:} networks, spatial ecology, co-occurrence, probability of interaction\\

\textbf{Words in the abstract:}

\textbf{Words in the main text:}

\textbf{Words in the legends:}

\textbf{Figures:}

\textbf{Tables:}

\textbf{References:}

\newpage
\doublespacing

%========================================================%
\section*{Abstract}

\newpage
%========================================================%
\section*{Introduction}

Ecology textbooks define the structure of an ecological community as
\emph{the pool of species occupying a given location at a given time, and
the way these species interact} (ref to Morin?). It is therefore surprising
that ecological research on the variation of community structure has focused
on species turnover (refs), almost entirely neglecting variation in the way
species interact (refs).

Network structure vary in space and time, metaweb as a large scale
synthesis. We don't know yet the mechanisms of variation

The metaweb concept is making its way through the network literature even
though it has never been formally and technically defined.
It is usually conceived as a network of interactions among species that
could potentially co-occur, and represented by a binary matrix, \emph{i.e.}
a probabilistic matrix in which the probability of observing an interaction
is either 0 or 1. As the metaweb is usually constructed by aggregating
observations, it assigns a probability of 0 to the interaction of species that
never co-occurred, and is unable to factor-out the effect of co-occurrence. The
main problem with that approach is that the co-occurrence of rare species is
extremely unlikely and thus most often appear as an absence of interactions
in the metaweb. A rarefaction analysis previously shown that interactions
accumulate with the addition of networks at a slower rate than species richness
\parencite{Poisot2012d} -- this is because interactions cannot accumulate
faster than the rate of species observations, and ultimately the rate of
co-occurrence observations. The only way to build a metaweb \emph{only}
based on observational data is to compensate for the inflation of 0 due
to lack of co-occurrence through manipulative experiments (REF). As this is
impracticable for all but the smallest organisms, predictive probabilistic
methods are needed.

No theory to explain and interpret the meaning of network variation in space.
Current interpretation of species turnover involves the effect of the
environment and stochasticity.

Objective: Propose a theoretical framework to understand and predict the
spatial and temporal variation in network structure.

\newpage
%========================================================%
\section*{A probabilistic representation of ecological interactions networks in space}

%% We need to be careful: ecological network / interaction network / etc.
%% I vote interaction networks, ecological network can be a spatial graph.
%% Also, since we're going to say "interaction" a lot, interaction network
%% makes sense. -- tp

Networks of species interactions vary in space and time because any given
pairwise interaction could either occur or not at any particular location.
Understanding this variation amounts to be able to determine the probability
that two species ($i$ and $j$) will interact at a given time/location ($y$);
the existence of an interaction in this situation is thus essentially a
stochastic variable, which we note $L_{ij,y}$. It takes a value of 1 where the
interaction occurs, and 0 elsewhere. We note $P(L_{ij,y})$ the probability
of this event. There are several factors that could impact the occurrence
of an interaction and we will describe them below. But ultimately, this
probability depends on the spatial and temporal scale of observation. As long
as the probability of a particular interaction is not null, the probability
of observing it at least once will become asymptotically closer to 1 when
the scale of observation increases.
% OK I like this paragraph, but we're REALLY NOT talking about factors or mechanisms here.
% This whole framework is top-down/inferential, so we should at least have a paragraph in the
% intro essentially summing up our preprint on network variation, and be clear on the
% fact that we don't touch on mechanisms.

The occurrence of an interaction requires the co-occurrence of species $i$ and
$j$. This argument might seem trivial at first, but the explicit consideration
of this condition in the probabilistic representation of ecological networks
will prove fundamental to understand their variation.
% We already make this point in betalink, should we summarize it briefly?
We thus define $X_{i,y}$
as a stochastic variable representing the occurrence of a species $i$ at
location $y$, and similarly $X_{ij,y}$ the co-occurrence of species $i$ and
$j$. The quantity we seek to understand is the probability of a joint event:

%-----------------
\begin{equation}
\text{P}(X_{i,y},X_{j,y},L_{ij,y},)
\end{equation}
%-----------------

Or simply said, the probability of observing both species $i$ and $j$, and
an interaction from $i$ to $j$. This probability could be further decomposed
in two parts using the product rule of probabilities:

%-----------------
\begin{equation}
P(X_{iy},X_{jy},L_{ijy})=P(L_{ijy}|X_{iy},X_{jy},E_y)P(X_{iy},X_{jy}|E_y)
\end{equation}
%-----------------

We will refer to the left term as the metaweb. It is a conditional
probability, representing the probability that an interaction occurs if
species $i$ and $j$ are co-occurring. The right term is the probability
of observing the two species co-occurring at location $y$. There are many
variants of the metaweb representing different hypotheses about the origin
of temporal and spatial variation in network structure (see the explicit
formulations at Table 1). First, the interaction could be deterministic
instead of probabilistic. In other words, $P(L_{ijy}) = 1$ if $X_{ijy} =
1$, and 0 otherwise. This representation of the metaweb is the one mostly
used so far, as soon as the species are found together they are assumed to
interact. It is also the only way to represent interactions when there is
not enough information available to evaluate the interaction probability. It
should be a reasonable approximation when the sampling and inference scales
are large enough and that the only variation of networks considered arises
from species distribution.

Ecological interactions could also depend on the environment. Although it
is not common to see a conditional representation of ecological interactions,
experimental studies of pairwise interactions revealing their sensitivity to
the environment are common (REF). For instance, it has been documented that the
predation risks of shorebirds do vary at the continental scale, from the south
to the north (REF). The effect of the environment on interactions propagate
up the community and influence network structure (REF). Here the environment
is considered in a very broad sense, as any factor potentially influencing the
probability of a pairwise interaction, provided that the species co-occur. It
thus includes both the biotic and the abiotic components of environment. We
note however that here the biotic environment includes organisms that are
not considered in the co-occurrence matrix. Including a biotic component to
the metaweb signifies that the pairwise interaction is conditional on higher
order interactions. An interaction modifier occurs for instance when the
predation risk by species $j$ might be impacted by a parasite $k$ changing
the behaviour of the prey $i$. We note that a conditional probability approach
could thus be used represent non-trophic interactions into ecological networks
(REF). This topic is however beyond the scope of the current paper.

There are also variants to the co-occurrence matrix. Akin to the metaweb,
co-occurrence could be conditional or not. The simplest model relates
co-occurrence probability directly to the environment. In this situation there
is no underlying assumption about the ecological processes responsible for
co-occurrence. Alternatively, the co-occurrence probability could be a function
of the environment because of shared ecological requirements by the two species.
Species are independently distributed, but co-occur more often that expected by
chance alone because they are found on the same environments. We call this model
later neutral because species are specifically responding to the environment but
are independently distributed. Co-occurrence is then simply obtained by
multiplying the result of two independent and specific species distribution
models (SDM).

Finally, the co-occurrence probability itself could be dependent on ecological
interactions. Direct pairwise interactions such as competition, facilitation and
predation have long been studied for their impact on co-distribution. Second and
higher order interactions (e.g. trophic cascade) could also impact
co-occurrence. There is however currently no general theory on the expected
co-occurrence in complex ecological networks. For instance, we do not know how
far co-occurrence is not-random when going along the chain of indirect
interactions. Berlow (2009) shown previously that almost only first and second
order interactions do matter in ecological networks, but we don't know for
co-distribution. We neither know what is the sensitivity to species richness: do
interactions tend to buffer each other? Generalizing knowledge acquired by the
study of small community modules will require future research.

\newpage
%========================================================%
\section*{Interpretation: the integrated niche}

Correctly describing the ecological niche is key to understand and predict
species distribution. Despite several attempts to refresh the conceptual
basis of what ecological niches are, ecologists haven't moved far past the
"n-dimensional hypervolume" formalism \parencite{Holt2009} introduced by
\textcite{Hutchinson1961}. Despite its intuitive interpretation and translation
into species distribution models \parencite{Blonder2014}, the concept has been
constantly criticized (Hardin, 1960; Peters, 1991; Chase2003; Silvertown, 2004;
Soberon, 2007) and several attempts have been made to expand and reinforce it.

Part of the problem surrounding the definition of the niche has been
clarified with the distinction between Eltonian and Grinnellian definitions
\parencite{Chase2003}. The Grinnellian dimension of the niche is the
effect of the environment on the demography of a species, while the Eltonian
dimension is the effect of a species on its environment \emph{sensu lato}. The
Grinnellian niche is the most intuitive one to apply and is the conceptual
backbone of species distribution models. The Eltonian niche is well known
by community ecologists, but is trickier to turn into predictive models
\parencite{Devictor2010a}. Nonetheless, the development of the niche model
of food web structure (Williams2000) and its parameterization (Williams2010;
Gravel2013) made it more operational, although it has yet to be applied to
more than trophic interactions.

\textcite{Chase2003} attempted this representation in their definition:
\textit{[The niche is] the joint description of the environmental conditions
that allow a species to satisfy its minimum requirements so that the birth rate
of a local population is equal or greater than its death rate along with the
set of per capita effects of that species on these environmental conditions.}
Their representation merges zero-net growth figuring the Grinnelian niche
(when do the population persist) with impact vectors figuring the Eltonian
niche (what is the population impact). While this representation has been very
influential in community ecology at the local scale, it remains impracticable
%%% WHY?
at the biogeographical one. The absence of any mathematical representation
of the niche that could easily be fit to ecological data perhaps explain why
biogeographers are still struggling to develop species distribution models
taking into account ecological interactions.

The key point to integrate dimensions of the niche is to represent the Eltonian niche into a Grinnelian space. 
- We do so by considering that the Eltonian niche is the hyper volume in the trait-space allowing an interaction. \\
- Doing so, we could project both niches in a plane and find the hypervolume where an interaction should occur (Fig. 2). \\
- This visual representation is parallel to the probabilistic definition of interaction probability. \\
- We propose that the metaweb is the Eltonian dimension of the niche, while the matrix of co-occurrence is the Grinnellian dimension. \\
- Feedbacks between dimensions occur through the inclusion of co-occurrence in the metaweb, and interactions in the co-occurrence matrix. \\
- This approach radically change the representation of the niche, putting species distribution and ecological interactions at the same level. \\
- Fitting the probabilistic model allows the evaluation of link distribution and species distribution models. \\
- Moreover, the integrated niche concept facilitates the formulation of species distribution models taking into account biotic interactions (see the section Applications) \\

% TO BE COMPLETED

\newpage
%========================================================%
\section*{Example: network structure in different habitats}

In this section we provide an analysis illustrating the framework with an
empirical dataset of host-parasitoid networks. Data come from the study of
Tylianakis(2007) on the impacts of habitat modifications to the network
structure. The data consists of 48 networks with 4090 recorded interactions. The
advantage of replicated host-parasitoid networks is that usually every
interaction is observed, not inferred from a stationary metaweb. It thus allows
to evaluate interaction probability and to factor out the effect of
co-occurrence. Five habitats were sampled along a gradient of habitat
modification: forest, abandoned coffee agroforest, coffee agroforest, pasture
and rice culture. The metaweb consists of 9 parasitoids and kleptoparasites
(Hymenoptera: Eulophidae, Ichneumonidae, Leucospidae, Megachilidae and
Chrysididae; Dyptera: Bombyliidae) of 33 species of bees and wasps (Hymenoptera:
Apidae, Megachilidae, Mutilidae, Pompilidae, Sphecidae, Vespidae). The metaweb
is illustrated at Fig. 2, along with an example of one iteration of the metaweb.

Tylianakis (2007) investigated whether habitat modification affects
the structure of these networks They found a significant impact of
the habitat on their structure, despite little variation in species
richness. Increasing habitat modification led to a higher parasitoid
to host species ratio and a paraistoid were also more specialized, thus
impacting considerably vulnerability. A closer inspection of the networks
revealed that intensive agricultural systems were dominated by a strong
interaction and a specialization of the most abundant parasitoid. Although
the discussion made clear that both the turnover in species composition and
the interaction probability changed with habitat modification, it was not
possible to partition these components.

We developed the \texttt{probaweb} package for R (REF) to fit
alternative formulations of the metaweb and the co-occurrence matrix
along an environmental gradient and run it to re-interpret the data of
Tylianaks (2007). The package provides a general interface facilitating
the development of different species and link distribution models. It is
also built to facilitate the interaction with the\texttt{mangal} database
of ecological interactions (REF). The first step consists of fitting a
probabilistic model from the observation of a pairwise interaction (binary)
and the environment (could be categorical or continuous) from the subset of
the data where the two species are co-occurring. In other words, it fits
the equation $P(L_{ijy}|X_{iy},X{jy},E_y)$ to the data where $X_{iy} = 1$
and $X_{iy} = 1$. Logistic regression was used and is currently programmed,
but alternative models could be used as well. The second steps consists of
fitting a probabilistic model for co-occurrence over the whole dataset,
$P(X_{iy},X{jy}|E_y)$, independently of the observation of an interaction. The
two probabilities are then multiplied to obtain the probability of observing
an interaction (Eq. 2). We used this probability to compute the likelihood of
each observation ($\zeta(\theta|D) = P(L_{ijy},X_{iy},X_{jy})$ if $L_{ijy}=1$
and $\zeta(\theta|D) = 1 - P(L_{ijy},X_{iy},X_{jy})$ otherwise). We then
after compare the models by their AIC.

We considered the gradient of habitat modification as a ordered categorical
variable and compared XX models (results are summarized at Table 2). Not
surprisingly the best model takes into account the effect of the environment on
both the metaweb and co-occurrence. What is most interesting are the comparisons
to the best model. First, we find that using a constant metaweb has a dramatic
impact on the fit of the model to the data (the AIC drops from X for model 1 to
X for model 2), indicating a strong effect of the environment on pairwise
interactions. Secondly, we find that the deterministic metaweb is the worst
model (model 3, AIC = ). This result indicate that the traditional approach to
consider that species interact as soon as they co-occur is definitely wrong.
Thirdly, we also find that using a constant co-occurrence does have a
significant impact on the model (the AIC drops to X, model 4), indicating there
is a non-random change in community composition with habitat modification. Taken
together, these two results better explain why network structured changed with
habitat modification, even though here we only used binary information about the
network structure. Another interesting result is that considering a neutral
co-occurrence did not impact much the fit of the model. The AIC drops to XX
with model 6, indicating that considering independent SDMs yields similar
networks over this environmental gradient. This means that for this particular
dataset, ecological interactions does not have a strong impact on species
distribution since; a strong dependence of parasitoids to the host for instance
would have a occurrence probability higher than expected by chance, while a
repulsion would have had the opposite.

An important output of this analysis is a more explicit representation of the
uncertainty in the evaluation of the metaweb. We find that among the XX pairs of
host and parasitoids, XX did not co-occur. There were therefore many forbidden
links based on co-occurrence. These might never occur in reality, but we do not
know without doing extra experiments. Therefore, any analysis of the structure
of the metaweb would be inappropriate without filling those gaps. In addition to
specific experiments, the gaps could be filled with a trait-based approach,
using phylogenies or with a null hypothesis (e.g. the interaction probability is
equal to connectance computed on the observed interactions).

It is also possible to obtain for each pairwise interaction an estimate of
the uncertainty. Not surprisingly, the confidence interval is usually high
for the estimation of a probability with a small sample size. The standard
error on the evaluation of the interaction probability is provided along
with the metaweb at Fig. 3. It reveals that the uncertainty is high for
most interactions, even if 48 networks were sampled. Such an approach could
be used to detect which pairwise interaction requires additional sampling
in order to reduce the uncertainty to a manageable level.

\newpage
%========================================================%
\section*{Applications}

\subsection*{Network descriptors}

\subsection*{Partitionning beta diversity}

\subsection*{Guidance for empirical studies}

Ecological networks are known to be extremely sparse, \emph{i.e.} having far
more absences of interactions that they have interactions. These absences of
interactions, however, can come from different sources. The fact that unequal
sampling at the local scale can affect our understanding of network structure
is well documented \parencite{Martinez1999}. However, in a spatial context,
some interactions may be undocumented because the species involved have
never been observed in co-occurence. Although these are reported as a lack
of interactions, in actuality we cannot make inference about them seeing
that they have never been observed: it is possible that this interaction
may happen should the two species co-occur. A second category of absences
of interactions are those that are reported after multiple observations of
species co-occurence. However, so as to have a confidence in the fact that the
probability of an interaction is low, extensive sampling (that is, several
co-occurences) is needed. Generally, our confidence that the interaction
is indeed impossible will increase when the number of observations of the
species pair. Seeing that this is essentialy a Bernoulli process (what is
the probability that the species will interact given their presence), the
breadth of the confidence interval is expected to saturate after a fixed
number of observations, which can be set as a treshold above which a species
paier has been observed "often enough".

\subsection*{Null model testing}

% Do you mean "Patterns-generating models testing"?

\subsection*{Species \& interactions distribution models}

\newpage
%========================================================%
\section*{Conclusion}

\begin{itemize}
\item New research agenda\\

\item List of new questions \\
\end{itemize}

\newpage
%========================================================%
\section*{Acknowledgements}
This is a contribution to the working groups \emph{Networks over ecological
gradients} (Santa Fe Institute) and \emph{Continental-scale variation of
ecological networks}. (Canadian Institute for Ecology and Evolution). DG
received financial support from NSERC and Canada Research Chair program. TP
is supported by a MELS-FQRNT post-doctoral fellowship.
\newpage

%========================================================%
\printbibliography

\newpage
%========================================================%
\section*{Figure legends}

%------------------------
\subsection*{Figure 1}
\textbf{Non-random sampling of the metaweb}. The sampling of the metaweb is illustrated with a local interaction network from the Tylianakis et al (2007) dataset. Here the metaweb is simply the number of observed interactions over the 48 networks. Arrows points to species that are present in the local network \#34 (represented on the right).

%------------------------
\subsection*{Figure 2}
\textbf{The integrated niche}. The Grinnellian niche is the set of environmental conditions
where the intrinsic growth rate $r$ is positive (axioms i, ii \& iii).
Contingencies (axiom iv) such as disturbances and stochastic extinctions, in
conjunction with limited dispersal (axiom v), restrict species distribution to
the conditions where the colonization rate $c$ is larger than extinction rate
$e$. The Eltonian niche on the other dimension is represented by the set of
traits allowing species to interact (axiom viii). The red species is a predator
with a trophic position $n$, feeding on species whose niche is within a certain
range around the niche optimum $c$. The integrated niche combines the effects of
the environment and ecological interactions. The central square represents the
area where the joint probability of observing interactions and co-occurrence is
positive.

%------------------------
\subsection*{Figure 3}
\textbf{Uncertainty in the evaluation of the metaweb}. The inferred metaweb for the XX environment is represented (left), along with the uncertainty in the evaluation of the interaction probability (right). Note that the standard error for probabilities is not symetric and thus only the upper bound is represented.

\newpage
%========================================================%
%------------------------
\subsection*{Figure 1}

\begin{figure}[ht!]
\centering\includegraphics[width=0.8\textwidth]{sampling}
\end{figure}

\newpage

%------------------------
\subsection*{Figure 2}

\begin{figure}[ht!]
\centering\includegraphics[width=0.8\textwidth]{niche}
\end{figure}

\newpage

%------------------------
\subsection*{Figure 3}

\begin{figure}[ht!]
\centering\includegraphics[width=0.8\textwidth]{niche}
\end{figure}

\newpage

%========================================================%

%------------------------
\begin{table*}[c]
\centering
\begin{tabular}{p{4cm}p{4cm}p{6cm}}
\hline
Name & Equation & Details \\
\hline
\textbf{Metaweb} &  & \\
Constant & $P(L_{ijy}|X_{iy},X_{jy})$ & Interaction probability is invariant to the environment \\
Conditional & $P(L_{ijy}|X_{iy},X_{jy},E_y)$ & Interaction probability is a function of the local environment \\
Deterministic & $P(L^*_{ijy}|X_{iy},X_{jy})$ & Interaction occurs whenever both species are present \\

\textbf{Co-occurrence} & &\\
Constant & $P(X_{iy},X_{jy})$ & Species distribution independent of $E$ \\
Conditional on $E$ & $P(X_{iy},X_{jy} |E_y)$ & Similar to a SDM applied to co-occurrence \\
Neutral & $P(X_{ix}|E_y)P(X_{jy}|E_y)$ & Independent SDMs fit to both species; could be independent of $E$ \\
Conditional on $L_y$ & $P(X_{iy},X_{jy} | L_y)$ & Could account for first and higher order interactions \\
\hline
\end{tabular}
\caption{List of different models}
\end{table*}

\newpage

%------------------------

\begin{sidewaystable}[c]
\centering
\begin{tabular}{lllllllll}
\hline
\textbf{Model} & \textbf{Metaweb} & & & \textbf{Co-occurrence} & & & \textbf{$L(H|D)$} & AIC\\
& Constant & Cond. on $E$ & Deterministic & Constant & Cond. on $E$ & Neutral & & \\
\hline
1. & X & & & & X & & 0 & 0 \\
2. & & X & & & X & & 0 & 0 \\
3. & & & X & & X & & 0 & 0 \\
4. & & X & & X & & & 0 & 0 \\
5. & & X & & & X & & 0 & 0 \\
6. & & X & & & & X & 0 & 0 \\
\hline
\end{tabular}
\caption{Model comparison with the host-parasitoid networks. The 48 networks were fitted to different models of interaction networks. Note that for the computation of the likelihood all null interaction probabilities, co-occurrences and the pairwise interactions without observed co-occurrences were removed.
}
\end{sidewaystable}

\end{document}
%========================================================%