### 2018-08-30 ms/ms_probaweb.tex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{color,lineno,setspace,multirow}
\usepackage{graphicx}
\usepackage[top=2.4cm,left=2.4cm,top=2.4cm,bottom=2.4cm,includefoot]{geometry}
\usepackage{natbib}
\usepackage{caption}
\usepackage{pdflscape}
\usepackage{hyperref}

\bibliographystyle{bes}

\begin{document}

\linenumbers
\modulolinenumbers[1]

\textbf{Title:} Bringing Elton and Grinnell together: a quantitative framework
to represent the biogeography of ecological interaction networks\\

\textbf{Authors:} Dominique Gravel$^{1,2,*}$, Benjamin Baiser$^{3}$, Jennifer
A. Dunne$^{4}$, Jens-Peter Kopelke$^{5}$, Neo D. Martinez$^{6}$, Tommi
Nyman$^{7}$, Timothée Poisot$^{2,8}$, Daniel B.
Stouffer$^{9}$, Jason M. Tylianakis$^{9,10}$, Spencer A. Wood$^{11}$, Tomas Roslin$^{12}$ \\

\textbf{1}: Canada Research Chair in Integrative Ecology. Département de
biologie, Université de Sherbrooke, 2500 Boulevard l'Université,
Sherbrooke (Québec). J1K 2R1\\
\textbf{2}: Québec Centre for Biodiversity Science\\
\textbf{3}: University of Florida, Department of Wildlife Ecology and Conservation, 110 Newins-Ziegler Hall, PO Box 110430, Gainesville, Fl. 32611-0430 \\
\textbf{4}: Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA.\\
\textbf{5}: Senckenberg Research Institute and Natural History Museum, Senckenberganlage 25, D-60325 Frankfurt am Main, Germany\\
\textbf{6}: Ecology and Evolutionary Biology Department, University of Arizona, P.O. Box 210088, Tucson, AZ 85721, USA\\
\textbf{7}: University of Eastern Finland, Department of Environmental and Biological Sciences, P.O. Box 111, FI-80101 Joensuu, Finland\\
\textbf{8}: Université de Montréal, Département des Sciences Biologiques, 90 Avenue Vincent d’Indy, Montréal, QC H2V3S9, Canada.\\
\textbf{9}: Centre for Integrative Ecology, School of Biological Sciences, University of Canterbury, Private Bag 4800, Christchurch, New Zealand\\
\textbf{10}: Department of Life Sciences, Imperial College London, Silwood Park Campus, 
Buckhurst Road, Ascot, Berkshire SL5 7PY, United Kingdom\\
\textbf{11}: University of Washington, School of Environmental and Forest Sciences,
Box 352100, Seattle, WA 98195, USA\\
\textbf{12}: Department of Ecology, Swedish University of Agricultural Sciences, Box 7044, 750 07 Uppsala, Sweden\\

%\textbf{Words in the abstract: 226} \\
%\textbf{Words in the main text: 7043} \\
%\textbf{Words in the legends: 502} \\

Authorship: DG, BB, JAD, NDM, TP, SAW, DBS and JMT contributed to the
concepts; JPK, TN and TR contributed the data; DG developped the probabilistic
approach and performed the analysis; DG, TP and TR wroted the first draft and
all of the authors edited the manuscript.

\newpage
\doublespacing

%=============================================================================%
\section*{Abstract}

Biogeography has traditionally focused on the spatial distribution and
abundance of species. Both are driven by the way species interact with one
another, but only recently community ecologists realized the need to document
their spatial and temporal variation. Here, we call for an integrated
approach, adopting the view that community structure is best represented as a
network of ecological interactions, and show how it translates to biogeography
questions. We propose that the ecological niche should encompass the effect of
the environment on species distribution (the Grinnellian dimension of the
niche) and on the ecological interactions among them (the Eltonian dimension).
Starting from this concept, we develop a quantitative theory to explain
turnover of interactions in space and time -- \emph{i.e.} a novel approach to
interaction distribution modeling. We apply this framework to host–parasite
interactions across Europe and find that two aspects of the environment
(temperature and precipitation) exert a strong imprint on species co-
occurrence, but not on species interactions. Even where species co-occur,
interaction proves to be stochastic rather than deterministic, adding to
variation in realized network structure. We also find that a large majority of
host-parasite pairs are never found together, thus precluding any inferences
regarding their probability to interact. This first attempt to explain
variation of network structure at large spatial scales opens new perspectives
at the interface of species distribution modeling and community ecology.

\textbf{Keywords:} networks, spatial ecology, co-occurrence, probability of interaction\\

\newpage

%=============================================================================%
\section*{Introduction}

Community ecology is \textit{the study of the interactions that determine the
distribution and abundance of organisms} \citep{Krebs2001}. Despite a general
consensus on this definition \citep{Scheiner2007a}, research on variation in
community structure (beta-diversity) has mostly focused on the spatial and
temporal turnover of species composition \citep{Anderson2011}. Such research
usually neglects variation in the way species interact with each other,
despite accumulating empirical evidence that is a major source of
diversity \citep{Poisot2015a}. Given this omission, it is perhaps not
surprising that biogeographers are still struggling to establish whether
interactions actually impacts the distribution of species at large spatial
scales \citep{Kissling2012}. An interaction is conceived as the direct effect
of the action of one species on the demography on another species; whether it
scales up and impact the (co-)distribution of species remains matter of debate
at the moment \citep{Wisz2012, GonzalezSalasar2013, Cazelles2016b,
Harris2016, Godsoe2017}. Treating interactions as fixed events nonetheless
neglects a large part of the complexity of empirical communities, and will
most likely deliver biased metrics \citep{Poisot2015c}. Recent attempts
at accounting for interactions in species distribution models
\citep{Pollock2014, Pelissier2013, Ovaskainen2017} have brought some
methodological advances, but are not sufficient for two reasons. First, these
techniques are still based on a `species-based' approach to communities, where
interactions are merely treated as fixed covariates affecting distribution.
Second, they fail to provide a conceptual step forward, both in their
treatment of interactions and in the quality of the predictions they make.

Network approaches offer a convenient representation of communities because
they simultaneously account for species composition and their interactions.
Species are represented as nodes, so that networks already encompass all the
information used by current approaches of species distribution modeling; in
addition, interactions are represented by links, so that networks provide
additional information on community structure. To date, studies of network
diversity have mostly been concerned with the distribution of interactions
within locations, and less so with variation among locations \citep{Dunne2006,
Bascompte2007, Ings2009, Kefi2012}. There is, however, ample evidence that
interaction networks vary in space and time \citep{Laliberte2010, Poisot2012,
Schleuning2012, Albouy2014, Poisot2016, Trojelsgaard2015}, even though there
is no common framework with which to generalize these results. Metacommunity
theory provides explanations for variation in the distribution of the nodes
\citep{Gravel2011c, Pillai2011,Cazelles2015}, but there are no such explanation
to the variation of node and link occurrences. Consequently, we need theory to
formalize these observations, as it is the only way towards fulfilling the
goal of community ecology: providing cogent predictions about, and
understanding of, the structure of ecological communities.

Given the historically different approaches to modelling the distributions of
species vs. interactions, there is a need to bring the two together. Here, we
offer an integrated approach to do so, adopting the view that community
structure is best represented as a network of ecological interactions. Based
on this idea, we propose a description of the ecological niche that
integrates the effect of the environment on species distribution and on the
ecological interactions among them. Building on this concept, we develop a
quantitative theory to explain turnover of interactions in space and time. We
first present the concept and then formalize it mathematically, using a
probabilistic model to represent the sampling of the regional pool of
interactions. At the level of species pairs, the statistical approach could be
conceived as an interaction distribution model. At the community level, the
approach provides a likelihood-based method to compare different hypotheses of
network turnover. As an illustrative example, we apply this novel framework to
a large data set on host-parasite interactions across Europe and find that two
aspects of the environment (temperature and precipitation) exert a strong
imprint on species co-occurrence, but not on species interactions. The network
structure changes systematically across the latitudinal gradient, with a peak
of connectance at intermediate latitudes.

%=============================================================================%
\section*{The two dimensions of community structure}

The problem of community assembly is often formulated as \textit{how are
species sampled from a regional pool to constitute a local community}
\citep{Gotzenberger2012}? This question could be rewritten to address the
problem of network assembly, as \textit{how do samples from a regional pool of
interactions constitute a local interaction network?} An illustration of this
problem for a food web is provided in Fig. 1. The regional pool of
interactions, the \emph{metaweb}, represents potential interactions among all
species that could be found in a given area. In this particular case, there
are 275 nodes, and 1173 links among the plants (52 nodes), herbivores (96
nodes), and parasitoids (127 nodes) from Northern Europe. An instance of a
local community is also illustrated, with 45 nodes and 93 interactions. Only
55.0\% of all potential interactions (plant-herbivore or herbivore parasitoid
combinations) are realized in this local network, revealing the stochastic
nature of ecological interactions. Our objective here is to provide theory to
explain the sampling of the regional pool of interactions, along with a
quantitative method to predict it. The problem could be formalized
sequentially by understanding first why only a fraction of the species co-occur locally and second why these species do or do not interact.

There are multiple causes of spatial turnover of species co-occurrence. The
first and most-studied driver is the effect of variation in the abiotic
environment on species performance. Combined with specific responses in
demography, it generates variation among sites by selecting the locally
fittest species \citep{Leibold2004a}. Stochasticity plays an additional role,
either because colonization and extinction events \citep{Hanski1999} are
inherently unpredictable or because strong non-linear feedbacks in community
dynamics generate alternative transients and equilibria \citep{Chase2007,
Vellend2014}. Interaction themselves may impact co-distribution, as
hypothesized long ago by \citet{Diamond1975}. Analyses of community turnover
are usually performed with data represented in a table with rows corresponding
to sites (or measurements) and columns to species. Metrics of beta diversity
quantify the variance of this community data \citep{Legendre2005}. Traditional
approaches rely on measures of dissimilarity among communities, such as the
Jaccard or Bray–Curtis indices. More recent approaches decompose total
variation of the community data into species and site contributions to beta
diversity \citep{Legendre2013}, and further partition it into dissimilarity
due to changes in species richness and dissimilarity due to actual species
turnover \citep{Baselga2010}. Even though all of these methods compare whole
lists of species among sites or measurements, they remain fundamentally
"species-based", since they report variation within columns. None of them
explicitly considers variation of associations (i.e., of pairs or higher-order
motifs – \citealt{Stouffer2007}). Some previous studies have considered how
species distribution could be influenced by the joint effects of the abiotic
and biotic environment \citep{Stephens2009, GonzalezSalasar2013, Cazelles2015,
Ovaskainen2017}, here we inverse the problem and describe how the distribution
of biotic interactions is influenced by species distribution and the
environment.

We are now getting a better understanding of interaction turnover. As
mentioned above, in the network approach to community structure, species and
interactions are represented by nodes and links, respectively. Associations
can also be represented by matrices in which entries represent the occurrence
or intensity of interactions among species (rows and columns). Network
complexity is then computed as the number of interactions (in the case of
binary networks) or interaction diversity (in the case of quantitative
networks, \citealt{Bersier2002}). Variability in community structure
consequently arises from the turnover of species composition, along with
turnover of interactions among pairs of species. The occurrence and intensity
of interactions could vary because of the environment, species abundance, and
higher-order ecological interactions \citep{Poisot2015a}. Variation in community
composition can be independent of variation of ecological
interactions, suggesting that species and interaction distribution may well
respond to different drivers \citep{Poisot2012}.

The "niche" is by far the dominant concept invoked to explain species
distributions and community assembly, from the local to the global scale.
Following \citet{Hutchinson1957}, the niche is viewed as the set of
environmental conditions allowing a population to establish and persist (see
also \citealt{Holt2009a}). In other words, the niche is the location in that
multidimensional space allowing a species to have a positive growth rate
when rare \citep{Godsoe2017}.  Community turnover arises as a result of
successive replacement of species along an environmental gradient, in
agreement with the Gleasonian view of communities \citep{Gleason1926}. The
concept is straightforward to put into practice with species distribution
models, as it maps naturally on available distributional and environmental
data. Consequently, a vast array of statistical tools have been developed to
implement it (e.g. BIOMOD \citealt{Thuiller2003}, MaxEnt
\citealt{Phillips2006}). It is however much harder to account for ecological
interactions within this approach \citep{Peterson2011}. As such, these
interactions are often viewed as externalities constraining or expanding the
range of environmental conditions required for a species to maintain a viable
population \citep{Pulliam2000, Soberon2007}.

Interestingly, the food web literature also has its own "niche model" to
position a species in a community \citep{Williams2000a}, generalized later to
other types of interaction networks \citep{Eklof2013}. The niche of a species
in this context represents the multidimensional space of all of its
interactions. Each species is characterized by a niche position, an optimum
and a range over three to five different niche axes \citep{Williams2000a,
Eklof2013}. The niche model of food web structure and its variants have
successfully explained the complexity of a variety of networks, from food webs
to plant-pollinator systems \citep{Allesina2008, Williams2010, Eklof2013}.
This conceptual framework is, however, limited to local communities, and does
not provide any explanation for the turnover of network structure along
environmental gradients.

%=============================================================================%
\section*{The integrated niche}

Despite several attempts to update the concept of the ecological niche,
ecologists have not moved far beyond the "n-dimensional hypervolume" defined
by \citet{Hutchinson1957}. Despite its intuitive interpretation and easy
translation into species distribution models \citep{Boulangeat2012,
Blonder2014}, the concept has been frequently criticized \citep{Hardin1960,
Peters1991, Silvertown2004}, and several attempts have been made to expand and
improve it \citep{Pulliam2000, Chase2003, Soberon2007, Holt2009a,
McInerny2012b}.

Part of the problem surrounding the niche concept has been clarified with the
distinction between Eltonian and Grinnellian definitions \citep{Chase2003}.
The Grinnellian dimension of the niche is the set of abiotic environmental conditions
required for a species to maintain a population in a location. The Grinnellian
niche is intuitive to apply, and constitutes the conceptual backbone of
species distribution models. The Eltonian niche, on the other hand, is 'the
place of a species in its biotic environment, its relations to food and
enemies'. While this aspect of the niche is well known by community
ecologists, it is trickier to turn into predictive models. Nonetheless, the
development of the niche model of food web structure \citep{Williams2000a} and
its parameterization using functional traits \citep{Gravel2013, Bartomeus2016}
made it more operational.

These perspectives are rather orthogonal to each other, and this has resulted
in considerable confusion in the literature \citep{McInerny2012a}.
\citealt{Chase2003} attempted to reconcile with the following definition:
\emph{"[The niche is] the joint description of the environmental conditions that
allow a species to satisfy its minimum requirements so that the birth rate of
a local population is equal to or greater than its death rate along with the
set of per capita effects of that species on these environmental conditions"}.
Their representation merges zero-net-growth isoclines delimiting the
Grinnellian niche ("when does the population persist?") with impact vectors
delimiting the Eltonian niche ("what is the per-capita impact?"). While this
representation has been very influential in local-scale community ecology (the
resource-ratio theory of coexistence, \citealt{Tilman1982}), it remains
impractical at larger spatial scales because of the difficulties in measuring
it. The absence of any mathematical representation of the niche that can be
easily fit to ecological data may explain why biogeographers are still
struggling to develop species distribution models that also consider
ecological interactions. Thus, a more integrative description of the niche
will be key to understand spatial and temporal turnover in community
structure.

We propose to integrate the two perspectives of the niche using a visual
representation of both components (Fig. 2). The underlying rationale is that,
in addition to the environmental constraints on demographic performance (Fig. 2,
top panel), any organism requires resources to meet its metabolic demands and
to sustain reproduction (Fig. 2, nodes in network of bottom panel). Abiotic
environmental axes are any non-consumable factors affecting the demographic
performance of an organism. Alternatively, the resource axes are traits of the
resources that allow interactions with the consumers. By definition of what an
interaction is, these axes also influence the growth rate. The niche can
therefore be viewed as the set of abiotic environmental conditions (the
Grinnellian component) along with the set of traits (the Eltonian component)
that allow a population to establish and to grow at a location. This visual
representation make operational the theoretical interpretation of the niche
and species ranges by \citet{Godsoe2017}. Accordingly, each
species can be characterized by an optimal position along both the
environmental (x-axis) and the trait (y-axis) plane. The integrated niche is
then the hypervolume where interactions can occur and where a population has a
positive growth rate.

This approach changes the representation of the niche, putting species
distributions and ecological interactions into the same formalism. Moreover,
it allows the limits of the niche axes to be independent of each other (as in
the example in Fig. 2), or to interact.  Some of these axes may not be
independent. For instance, the optimal prey size for predatory fishes could
decline with increasing temperature \citep{Gibert2014}, which would make diet
boundaries functions of the environment. Alternatively, we could also consider
that the growth rate of the predator changes with the size of its prey items,
thereby altering the environmental boundaries and changing the shape of the
niche in the middle of this visual representation. It is also important that
this conceptual representation depicts a reality that could be highly
dimensional and sometimes difficult to represent statistically
\citep{Clark2007}.

%========================================================%
\section*{A probabilistic representation of interaction networks in space}

We now formalize the integrated niche with a probabilistic approach to
interactions and distributions. In particular, we seek to represent the
probability that an interaction between species $i$ and $j$ occurs at location
$y$. We define $L_{ijy}$ as a stochastic process taking a value of 1 when an
interaction occurs and a value of 0 when it does not, and focus on the
probability that this event occurs, $P(L_{ijy} = 1)$, over a given sampling
interval and space. We note that the occurrence of an interaction is dependent
on the co-occurrence of species $i$ and $j$. This argument might seem trivial
at first, but the explicit consideration of this condition in the
probabilistic representation of ecological interactions will prove
instrumental to understanding their variation. We note this phenomenon we aim
to describe differs fundamentally from the effect that interactions do have on
co-occurrence \citep{Cazelles2016}. We similarly define $X_{iy}$ as a
stochastic process representing the occurrence of species $i$ at location
$y$, taking a value of 1 when the species occurs and a value of 0 when it is
absent. The quantity we seek to understand is the probability of a joint
event, conditional on the set of environmental conditions $E_y$:

%-----------------
\begin{equation}
	\text{P}(X_{iy},X_{jy},L_{ijy}\lvert E_y)
\end{equation}
%-----------------

Or simply said, the probability of observing both species $i$ and $j$ plus an
interaction between $i$ and $j$ given the conditions $E_y$ at location $y$.
This probability could be decomposed into two parts using the product rule of
probabilities:

%-----------------
\begin{equation}
	P(X_{iy},X_{jy},L_{ijy}\lvert E_y)=P(X_{iy},X_{jy}\lvert E_y)P(L_{ijy}\lvert X_{iy},X_{jy},E_y)
\end{equation}
%-----------------

The first term on the right-hand side of the equation is the probability of
observing the two species co-occurring at location $y$. It corresponds to the
Grinnellian dimension of the niche. The second term represents the probability
that an interaction occurs between species $i$ and $j$, given that they are
co-occurring. This predicate can be refined using information on trait
distribution and trait matching rules . For brevity, we have here avoided
specific references to traits and considered that the interactions are
provided in the data. But one could develop the model further to integrate
such traits, inspired by the framework developed by Gravel et al. 2016. They
proposed to represent the occurence of an interaction between species as a
joint probability event $P(L_{ij}, \mathbf{T}_i, \mathbf{T}_j\lvert E)$, where
$\mathbf{T}_i$ and $\mathbf{T}_j$ are trait vectors for species $i$ and $j$
respectively. In order to develop a trait-based representation of the spatial
variation of interactions, one may thereafter consider how these traits are
related to the environment and how they co-occur. Above, we referred to this
entity as the "metaweb" and note that it corresponds to the Eltonian dimension
of the niche. Below, we will see how this formalism can be directly fit to
empirical data, restricting ourselves to occurrences only and refering to the
above citations for trait-matching and trait-environment relationships. But
before turning to an application, we will discuss the interpretation of
different variants of these two terms.

\subsection*{Variants of co-occurrence}

There are several variants to the co-occurrence probability, representing
different hypotheses concerning spatial variation in network structure (see
the explicit formulations in Table 1). The simplest model relates the
probability of co-occurrence directly to the environment,
$P(X_{iy},X_{jy}\lvert E_y)$. In this situation, there are no underlying assumptions
about the ecological processes responsible for co-occurrence. Spatial associations between species could arise
because interactions constrain distribution, where in such case the co-occurrence would be conditional on $L$, or, alternatively, because of
environmental requirements shared between $i$ and $j$ \citep{Pollock2014,
Cazelles2016}. In the former case, species are not independent of each other
and the conditional occurrence must be accounted for explicitly,
$P(X_{iy},X_{jy} \lvert E_y)=P(X_{iy}\lvert E_y,X_{jy})P(X_{jy}\lvert E_y)$. In the latter case,
species are independent, and only the marginal occurrence must be accounted
for, $P( X_{ijy}\lvert E_y)=P(X_{iy} \lvert E_y)P(X_{jy} \lvert E_y)$.

The co-occurrence probability itself could depend on ecological interactions.
This should be viewed as the realized component of the niche (i.e. the
distribution when accounting for species interactions). Direct pairwise
interactions such as competition, facilitation, and predation have long been
studied for their impact on co-distribution (e.g. \citealt{Diamond1975,
Connor1979, Gotelli2000}). Second- and higher-order interactions (e.g. trophic
cascades) could also affect co-occurrence \citep{Harris2016,
Staniczenko2017}. Co-occurrence of multiple species embedded in ecological
networks is a topic of its own, however, and is influenced by both network
topology and species richness \citep{Cazelles2016}. Not only direct
interactions influence co-occurrence, but indirect interactions do as well
(e.g. plant species sharing an herbivore, or herbivores sharing parasitoids,
could repel each other in space \citep{Holt1993}). The impact of direct
interactions and first-order indirect interactions on co-occurrence tends to
vanish with increasing species richness in the community \citep{Cazelles2016}.
Further, co-occurrence is also influenced by the covariance of interacting
species to an environmental gradient \citep{Cazelles2015}.

The interpretation that interactions may impact co-occurrence, and that the
realization of an interaction requires co-occurrence are not mutually
exclusive. But since here the object is to describe the variation in the
occurrence of interactions, which absolutely requires co-occurrence, we will
give attention to this part of the problem. Relating co-occurrence to the
structure of interaction networks is a problem on its own (and vice-versa). We
will therefore focus here on the variation of interactions and not on their
distribution, and leave this specific issue for the Perspectives section and
future research.

\subsection*{Variants of the metaweb}

There are also variants of the metaweb. First, most documented metawebs have
thus far considered ecological interactions to be deterministic, rather than
probabilistic (e.g. \citealt{Havens1992, Wood2015}). Species are assumed to
interact whenever they are found together in a location, independent of their
local abundance and the local environment. In other words,
$P(L_{ijy}\lvert X_{iy}=1,X_{jy}=1) = 1$. This approach might be a reasonable
approximation if the spatial or temporal scale of sampling and inference is so
large that the probability of observing at least one interaction converges to
unity. In this scenario, network variation arises solely from species
distributions.

Second, ecological interactions could also vary with the environment, so that
$P(L_{ijy}\lvert E_y)$. Although it is rare to see a conditional representation of
pairwise ecological interactions, experimental studies have frequently
revealed interactions to be sensitive to the environment. For instance,
\citet{Mckinnon2010} showed that predation risks of shorebirds vary at the
continental scale, decreasing from the south to the north. It is also common
to see increasing top-down control with temperature (e.g. \citealt{Shurin2012,
Gray2016}). Effects of the environment on interactions \citep{Gibert2014} also
propagate up the community and influence network structure
\citep{Tylianakis2007, Woodward2010, Petchey2010a}.

%========================================================%

\section*{Application: continental-scale variation of host-parasite community structure}

We now turn to an illustration of our theory with the analysis of an empirical
dataset of host-parasite networks sampled throughout the south-north
environmental gradient in continental Europe \citep{Kopelke2017}. Our
objective here is to illustrate potential applications of the approach
outlined above, rather than to specifically describe the biogeography
of this particular community. The focal system consists of local food webs of
willows (genus \textit{Salix}), their galling insects, and the natural enemies
(parasitoids and inquilines) of gallers. We ask:
i) how much does network structure vary across the gradient, and ii) what is
the primary driver of network turnover across the gradient?

\subsection*{Data}

Communities of willows, gallers, and parasitoids are species-rich and widely
distributed, with pronounced variation in community composition across space.
The genus \textit{Salix} includes over 400 species, most of which are shrubs
or small trees \citep{Argus1997}, and is common in moist habitats across the
Northern Hemisphere \citep{Skvortsov1999}. Willows support a highly diverse
community of herbivorous insects, with one of the main herbivore groups being
gall-inducing sawflies (Hymenoptera: Tenthredinidae: Nematinae: Euurina
\citep{Kopelke1999}. Gall formation is induced by sawfly females during
oviposition, and includes marked manipulation of host-plant chemistry by the
galler \citep{Nyman2000}. The enemy community of the gallers includes nearly
100 species belonging to 17 insect families of four orders
\citep{Kopelke2003}. These encompass two main types: inquiline larvae
(Coleoptera, Lepidoptera, Diptera, and Hymenoptera) feed primarily on gall
tissue, but typically kill the galler larva in the process, while parasitoid
larvae (representing many families in Hymenoptera) kill the galler larvae by
direct feeding \citep{Kopelke2003}.

Local realizations of the willow-galler-parasitoid network were reconstructed
from community samples collected between 1982 and 2010. During this period,
willow galls were collected at 370 sites across Central and Northern Europe.
In total, 52, 96 and 127 \textit{Salix}, galler and parasitoid and inquiline
taxa were distinguished, respectively. The strength of this dataset is that
observations were observed in situ, rather than inferred from expert knowledge
or other sources of information, thereby allowing the analysis of their
spatial variation. The drawback is that, because many species are rare, some
of the pairs of species may have been observed co-occurring only a few times,
or never. As a consequence, despite the extent of the sampling, there is
significant uncertainty in the quantification of some of the links. The above
described methodology explicitly aims at describing this uncertainty, and
research should be done to develop methods to reduce it (see Conclusion).

The current study represents the first analysis of the full data set
from a spatial perspective. Full details about data collection and species
identification are provided in \citep{Kopelke2017}. Annual mean temperature
and precipitation were obtained from WorldClim using the R package dismo
\citep{Hijmans2015} and GPS coordinates of the sampling locations. While other
covariates could have also been considered to represent the high
dimensionality of the niche, these two variables are likely representative of
the most important axes of European climate, and are also more easily
interpretable than reduced variables obtained, for example, by principal
component analysis.

\subsection*{Analysis}

Computing the probability of observing an interaction involves fitting a set
of binomial models and collecting their estimated probabilities. For the sake
of illustration, we considered generalized linear models – although more
flexible fitting algorithms (e.g. GAM or Random Forest) could equally well be
used, as long as the algorithm can estimate the probability for each
observation. The data consist of a simple (albeit large and sparse) table with
the observation of each species, $X_{iy}$ and $X_{jy}$, their co-occurrence,
$X_{ijy}$, the observation of an interaction $L_{ijy}$, and environmental co-
variates $E_y$. Thus, there is one row per pair of species per site. We
considered that an absence of a record of an interaction between co-occurring
species at a site means a true absence (see below for a discussion on this
issue).

We compared three models for the co-occurrence probability. The first one
directly represents the co-occurrence probability conditional on the local
environment, $P(X_{iy},X_{jy}\lvert E_y)$ (models are listed in Tables 1 and 2).
Hence, this model makes no assumption about the mechanisms driving co-
occurrence for any given environment, and instead uses the information
directly available in the data. It thereby indirectly accounts for the effect
of interactions on co-occurrence, if there are any. The second model considers
independent occurrence of species. In this case, we independently fit
$P(X_{iy} \lvert E_y)$ and $P(X_{jy} \lvert E_y)$, and we then take their product to
derive the probability of co-occurrence. This model should be viewed as a null
hypothesis with respect to the first model, since a comparison between the
respective models will reveal if there is significant spatial association of
the two species beyond a joint response to the shared environment
\citep{Cazelles2016}. Finally, the third model assumes that the probability of
co-occurrence is independent of the environment and thus constant throughout
the landscape. In other words, $P(X_{iy},X_{jy})$ is obtained by simply
counting the number of observed co-occurrences divided by the total number of
observations. Thus, the comparison between the first and third model allows us
to test the hypothesis that co-occurrence is conditional on the environment.
Whenever the environment was included as a covariate in the GLM, we considered
a second-order polynomial term for both temperature and precipitation in
order to account for optima in environmental conditions. There are
consequently five parameters for the first model when fitting a given pair of
species, 10 parameters for the second, and only one for the third model.

Following the same logic, we compared three models of the interaction
probability. The first model conditions the interaction probability on the
local environmental variables, $P(L_{ijy}\lvert X_{iy},X_{jy},E_y)$. Consequently,
the model was fit to the subset of the data where the two species co-occur.
The second model fits the interaction probability independently of the local
environmental variables, $P(L_{ijy}\lvert X_{iy},X_{jy})$. It corresponds to the
number of times the two species were observed to interact when co-occurring,
divided by the number of times that they co-occurred. The third model is an
extreme case evaluated only to test the hypothesis that if two species are
found to interact at least once, then they should interact whenever they co-occur, $P(L_{ijy}\lvert X_{iy},X_{jy})=1$. While not necessarily realistic, this
model tests an assumption commonly invoked in the representation of local
networks from the knowledge of a deterministic metaweb. It represents
potential interactions, rather than realized ones. There are consequently five
parameters for the first model, a single parameter for the second model and no
parameter to evaluate for the third model (where the interaction probability
is fixed by the hypothesis).

We fitted the different models to each pair of species and recorded the
predicted probabilities. The joint probability $P(L_{ijy},X_{iy},X_{jy})$ was
then computed from Eq. 2, and the likelihood of each observation $D_{ijy}$ given the model $\theta$ was computed
as $\mathcal{L}(D_{ijy}\lvert \theta)=P(L_{ij},X_{iy},X_{jy})$ if an
interaction was observed, and as $\mathcal{L}(D_{ijy}\lvert \theta) =
1-P(L_{ijy}, X_{iy}, X_{jy})$ if no interaction was observed. The log-likelihood was summed over the entire dataset to compare the different models
by AIC. We therefore evaluate the likelihood of all local networks, given the
model. Not surprisingly, there was a very large number of species pairs for
which this model could not be computed, as they simply never co-occurred. For
these pairs, we have no information of the interaction probability, and they
were consequently removed from the analysis. The log-likelihood reported
across the entire dataset was therefore summed over all pairs of species observed to co-occur at least once. Interactions between the first (\textit{Salix}) and
second (gallers) trophic layers and those between the second and third
(parasitoids) were considered separately. Finally, we used the full model (in
which both co-occurrence and the interaction are conditional on the
environment) to interpolate species distributions and interaction
probabilities across the entire European continent. We reconstructed the
expected network for each location in a 1 X 1 km grid and computed the
probabilistic connectance following \citet{Poisot2015c}.

All of the data are openly available in the database \emph{mangal}
\citep{Poisot2015b} and all R scripts for running the analyses, are provided
in the github repository
%\hyperref[https://github.com/DominiqueGravel/ms_probaweb]{https://github.com/DominiqueGravel/ms_probaweb}.
\url{https://github.com/DominiqueGravel/ms_probaweb}.

\subsection*{Results}

Despite the extensive sampling, many pairs of species were observed to co-
occur only a few times. This made it difficult to evaluate interaction
probabilities with any reasonable confidence. Thus, we start with an example
of a single pair of species selected because of its high number of co-
occurrences ($N_{ij}=38$): the leaf folder \textit{Phyllocolpa prussica} and
the parasidoid \textit{Chrysocharis elongata}. These two fairly abundant
species were observed $N_i=49$ and $N_j=121$ times, respectively, across the
370 sites, and they were found to interact with a marginal probability
$P(L_{ij})=0.55$, which means they interacted at 21 different locations. Here,
a comparison of model fit (Table 2) reveals that conditioning the interaction
probability on local environmental conditions adds no explanatory power beyond
a model assuming the same probability of interaction anywhere in space (Model
2 vs Model 3). Moreover, when the two species co-occur, the occurrence of the
interaction was insensitive to the environment (Model 2 vs Model 3).
Alternatively, climatic variables significantly impacted co-occurrence (Model
3 vs Model 4). The independent model performed worse than the non-random co-
occurrence model (Model 3 vs Model 6). The full model revealed that the
greatest interaction probability occurred at intermediate temperature and
precipitation, simply because this is where the two species most frequently
co-occur (Fig. 3). The probabilities of co-occurrence and interaction can be
represented in space, where we find that the highest interaction probability
occurred in Central Europe (Fig. 4).

We evaluated each model for all pairs of species in order to better understand
the large-scale drivers of network turnover. The results were highly
consistent among trophic levels (\textit{Salix}-gallers and
gallers-parasitoids; Table 3), despite large variability in the fit
of the models to the different pairs of species. This variability was
particularly important for the models of interactions given co-occurrence.
Across all pairs of species, the conditional representation of interactions
performed (Model 2) better than the marginal one (Model 1); that is,
interactions did not occur systematically whenever the two species were found
co-occurring. Hence, in addition to species turnover, the stochastic nature of
interactions contributes to network variability. In total, we recorded 1,173
pairs of interactions, only 290 of which occurred more than five times. Out of
these 290 interactions, 143 were systematically detected whenever the two
species co-occurred. In the instances when species co-occurred, the two
environmental variables considered proved relatively poor predictors of their
interactions (Model 2 vs Model 3). Not surprisingly, for both types of
interactions (\textit{Salix}-galler and galler-parasitoid), the log-likelihood
increased when the environment was considered. However, the extra number of
parameters exceeded the gain in log-likelihood and inflated AIC. Therefore,
the most parsimonious model excluded the effect of the environment. On the
basis of log-likelihood only, co-occurrence was non-independent for both
\textit{Salix}-galler and galler-parasitoid interactions. Thus, according to
AIC, the best model was the one of non-random co-occurrence (Model 3 vs Model
6) for both types of interactions.

% The approach we present not only has implications for understanding the
% biogeography of pairwise interactions and interaction networks, but also for
% evaluating the quality of metawebs. We investigated the reliability of the
% estimated metaweb across the entire dataset wtih summary statistics of species
% co-occurrence. As mentioned above, across the 17,184 potential pairs of
% species, only 1,173 pairs interacted in at least a single location, yielding a
% connectance of 0.068. However, only 4,459 pairs of species were found co-
% occurring at least once across all locations. There are consequently 12,725
% gaps of information in the metaweb (74.1\% - see Fig. 5). As we cannot know
% whether the non-co-occurring species would indeed interact if found
% together, a more appropriate estimate of connectance would be
% $C=1173/4459=0.263$. This result reveals that the evaluation of the sampling
% quality of ecological networks is a problem on its own and well worth further
% attention.

Once we had selected the best model based on AIC (Model 3, Table 2), we used
it to reconstruct the expected species richness, along with the most likely
network for each location. Using this approach, we mapped the expected
distribution of network properties across Europe (Fig. 5). For simplicity, we
chose to consider connectance as our descriptor of network configuration, as
this metric can be easily computed from probabilistic networks
\citep{Poisot2015c} and is also a good proxy for many other network properties
\citep{Poisot2014}. Overall, we found a peak in \textit{Salix}, gallers and
parasitoid diversity in Northern Europe. The expected number of interactions
roughly followed the distribution of species richness, but accumulated at a
rate different from species numbers. Connectance likewise peaked in Northern
Europe (Fig. 5).

%========================================================%
\section*{Interpretation}

We have proposed that the representation of community structure and its
variation in space is best captured by the formalism of ecological networks,
as both the distribution of species and their interspecific interactions can
then be accounted for. We consequently revised the niche concept in order to
integrate its abiotic and biotic components. This integrated niche was
represented visually with an ordination of species into an environmental space
and a trait space. The fundamental niche of a species is represented as the
set of environmental conditions and resources that allow a species to
establish in a location, thereby integrating the Eltonian and the Grinnellian
components of the niche. We then translated the concept mathematically by
investigating the probability of the joint occurrences of species and their
interaction, which should be interpreted as an interaction distribution model.
We used this approach to characterize the turnover of the structure of
ecological interactions in a species-rich tri-trophic network across Western
Europe, finding that the primary driver of network variation is the turnover
in species composition.

%To
%our knowledge, this is the first continental-wide analysis of the drivers of
%network structure from empirical data on the occurrence of interactions (see
%\citealt{Baiser2012, Albouy2014, Poisot2016}).

Applying the framework to our large data set on host-parasite interactions
across Europe revealed key features in the interaction between \textit{Salix}
taxa, their herbivores, and the natural enemies of these herbivores.
Consistent with a general increase in the diversity of \textit{Salix} towards
boreal areas \citep{Cronk2015}, overall species richness of the networks
increased towards the north. The distribution of \textit{Salix} species
richness largely matched those of gallers and parasitoids. These observations
within Europe are also matched by the ones found at a global scale for
\textit{Salix} \citep{Argus1997, Cronk2015, Wu2015} and sawflies
\citep{Kouki1994, Kouki1999}. Species richness in a common group of parasitic
wasps, the Ichneumonidae, was originally presumed to show a similar "reversed
latitudinal gradient", but this observation has been recently challenged by
findings of rather high ichneumonid diversity in the tropics
\citep{Veijalainen2013}. Nevertheless, ichneumonid subfamilies specifically
associated with sawflies (Ctenopelmatinae, Tryphoninae) are clearly less
diverse in the south.

Exactly what processes are responsible for the distribution of species
richness at different trophic levels is yet to be established (but see e.g.
\citealt{Roininen2005, Nyman2010, Leppanen2014}), but as a net outcome of
different latitudinal trends across trophic levels, the distribution of co-occurrence and therefore of potential interactions differed between the first
and second layers of feeding links. The correlation between expected
\textit{Salix} and galler richness was 0.73, while it was 0.58 between
gallers and their parasitoids. Therefore, the ratio of herbivores to
\textit{Salix} species is essentially constant across Europe, whereas each
herbivore species is potentially attacked by a richer enemy community higher
latitudes. Consequently, overall connectance peaks in Northern Europe (Fig.
5).

In terms of species interacting with each other, our analysis suggests that
the environment leaves a detectable imprint on species co-occurrence, but only
a slight mark on the occurrence of realized links among species in a specific
place: the probability of finding a given combination of species at a higher
and a lower trophic level at the same site was clearly affected by the
environment, whereas the probability of observing an interaction between the
two was, given co-occurrence, not significantly so. The interactions were  highly
uncertain, as only few pairs of species were systematically interacting when
co-occurring but most not. This applies to the example species
\textit{Phyllocolpa prussica} and \textit{Chrysocharis elongata} (Figs. 3 and
4), but also to all species pairs more generally. For the example species
pair, the full model revealed that the joint probability event (interaction
and co-occurrence) peaks at intermediate temperature and precipitation, simply
because this is where the two species co-occur most often. This does not imply
that species will always interact when they meet – although this is a basic
assumption in most documented metawebs to date (e.g. \citealt{Havens1992,
Wood2015}). Rather, an interaction is better represented as a stochastic
process whose probability is also influenced by the probability with which
species co-occur. What we cannot reliably know is how this stochasticity
splits into different sources of uncertainty – i.e., the extent to which a
species at the higher trophic level runs into a species at the lower level co-occurring at the site, the extent to which this interaction is detected by an
observer collecting a finite sample, or simply the uncertainty arising from
incomplete description of a highly dimensional niche. Future work will be
required to document the relative importance of these sources of uncertainty
in the occurrence of interactions.

%========================================================%
\section*{Perspectives}

Evidence that the structure of ecological networks does vary across habitats
(e.g. \citealt{Tylianakis2007}), over environmental gradients
\citep{Lurgi2012} and in time \citep{Simanonok2014} is accumulating
rapidly. It is not clear, however, to what extent the turnover of network
structure is driven by a systematic change in species composition or of
pairwise interactions \citep{Poisot2012, Poisot2015a}. Our model comparison of
host-parasite interactions revealed that most of the turnover is driven by
species-specific responses to the environment, impacting species richness, and
that co-occurrence was mostly independent. Further, the occurrence of interactions
among host and parasite is highly stochastic even when both are present, and
not predictable by the variables considered by us. We know that interactions
vary with the environment in other systems, for instance, herbivory
\citep{Shurin2012, Baskett2018} and predation \citep{Mckinnon2010,
Legagneux2014} are often found to increase with temperature, resulting in
spatial variation of trophic cascades \citep{Gray2016}. What remains unclear,
however, is the extent to which such variation is driven by a turnover of
species composition along gradients, or a turnover of the interactions. Here
we found that interactions vary substantially but non-predictably along the
annual temperature and the precipitation gradient. The lack of detectable
signal may be due to our choice of covariates. Indeed, a previous study focusing on a subset of the system studied here identified habitat characteristics as the primary drivers of
interactions \citep{Nyman2015}. New investigations with other systems will
thus be required to challenge this result. Under all circumstances,
documenting the relationship between the environment and the occurrence of
interactions at continental scales is critical for understanding how large-
scale variation of trophic regulation influences community dynamics and
ecosystem functioning \citep{Harfoot2014}.

The framework we provide complements methods to compute network beta-diversity
presented by \citealt{Poisot2012} and \citealt{Poisot2018}. The total network
turnover is partitioned into interaction turnover and species turnover – which
in our approach would correspond to the three terms of Eq. 2, respectively.
Importantly, one could easily derive the different $\beta$ of
\citealt{Poisot2012} using the fitted probabilities. For instance, for a pair
of species $i$ and $j$ and sites $x$ and $y$, the expectation for the fraction
$a$ for the interactions would be
$P(L_{ijx}\lvert X_{ix},X_{ijx})P(L_{ijx}\lvert X_{ix},X_{ijx})$. The same could be
computed for fractions $b$, and therefore the expected beta-diversity for an
entire network could be recomputed this way. The novel feature of the approach
presented here is then that the different contributions to network turnover
can be attributed to different hypotheses (as performed here), in order to
evaluate for instance the role of environmental heterogeneity or species co-
distribution on network turnover.

We restricted our analysis to the effect of co-occurrence on ecological
interactions, neglecting the inverse of the problem. We did not investigate in
depth the drivers of co-occurrence and simply took it for granted from the
data. Co-occurrence was indeed different from the expectation of independent
species distributions. It thus begs the question of whether, once
environmental effects on species-specific distribution have been accounted
for, interactions come with significant effects on co-occurrence? We could
rephrase this problem by asking whether the fundamental niche differs from the
realized niche, and how this applies to our framework. For example, we have
considered above simply the co-occurrence probability, $P(X_{iy},X_{jy}\lvert E_y)$,
which could be expanded as $P(X_{iy}\lvert X_{jy},E_y) P(X_{jy}\lvert  E_y )$. After some
re-arrangement of Eq. 2, the marginal occurrence probability, $P(X_{jy}\lvert E_y
)$, could be considered as a species distribution model taking into account
the interaction between these species (in the same spirit as done by Gonzalez-
Salazar et al. 2013). This derivation would however critically depend on a
strong \emph{a priori} expectation of the conditional probability of observing
a species given the distribution of the other species. This assumption seems
reasonable for some situations, such as a parasitoid species that requires a
host to develop. On the other hand, \citet{Cazelles2016b} found that the
strength of this association is often rather weak if not inexistant (for
instance, with the example pair presented in Table 1). The lack of an
association could simply arise when the parasitoid is generalist enough that
it is not constrained to track the distribution of any single/given host
\citep{Cazelles2015}.

The approach we presented could easily be expanded to account for traits, in
order to derive a complete and more mechanistic representation of the niche.
In this context, both the Eltonian and Grinnellian dimensions of the niche
could be represented. As mentioned earlier, \citealt{Gravel2016} proposed to
represent the occurrence of interactions based on some trait-matching rules
(given by a function of the type $P(L_{ij})\lvert \mathbf{T}_i, \mathbf{T}_j, E)$)
and the co-distribution of traits (corresponding to $P(\mathbf{T}_i,
\mathbf{T}_j\lvert E)$). The key  insight from the formalism proposed by
\citealt{Gravel2016} is that the co-distribution of traits will impact
variation in network structure. As a result, it was found that network
properties often co-vary with functional structure and functional diversity
\citealp{Laigle2018}. The next step in the development of this framework will
be to develop the models for the co-distribution of traits, for instance
following Ovaskainen et al. 2017. This may facilitate the accessibility and
the applicability of the framework, since the basic unit of analysis is no
longer the species, but rather a combination of traits, thereby facilitating
the statistical analysis by sharing information among species.

At present, there is only indirect support for the hypothesis that interacting
species are conditionally distributed, but this possibility should be the
topic of theoretical investigation with dynamical metacommunity models
\citep{Cazelles2015} and empirical hypothesis testing. The impact of
ecological interactions on the distribution of co-occurrence has been the
topic of many publications since \citet{Diamond1975} seminal study on
competition and "checkerboard" distribution, but pairwise approaches have only
recently received attention \citep{Veech2013}. Whether two interacting species
are more closely associated in space remains unclear, since most approaches
based on null models consider community-level metrics (e.g.
\citealt{Gotelli2000}), such as the C-score, thereby making it hard to
evaluate if specific interactions do indeed affect co-occurrence. The
expansion of the framework we describe to account for the difference between
the realized and the fundamental niche will therefore require further
investigation of the impact of interactions on co-occurrence.

Ecological networks are known to be extremely sparse, \emph{i.e.} they have
far more absences than presences of interactions. Absences of interactions,
however, can come from different sources. The fact that unequal sampling at
the local scale can affect our understanding of network structure is well
documented \citep{Martinez1999}. In a spatial context, however, some
interactions may be undocumented simply because the species involved have
never been observed to co-occur. Although these cases are reported as a lack
of interactions, in actuality we cannot make any reliable inference from them:
since the species have never been observed together, it remains possible that
they would interact if they did. A fundamentally different category of
absences of interactions are then those reported after multiple observations
of species co-occurence. Thus, to gain confidence that the probability of an
interaction is low, extensive sampling (that is, several records of co-
occurence) is needed. Generally, our confidence that the interaction is indeed
impossible will increase with the number of observations of the species pair.
Seeing that this is essentially a Bernoulli process (the probability that the
species will interact given their presence), the breadth of the confidence
interval is expected to saturate after a fixed number of observations, which
can be set as a threshold above which a species pair has finally been observed
"often enough". For instance, if two species are observed together at $k=10$
locations but are never found interacting ($n=0$), this means that the maximum
likelihood estimate for the probability of this binomial distribution is
$p=0$. Using the Clopper-Pearson method to compute the 95\% confidence
interval, we find that the real value of the probability of the interaction
occuring is somewhere between 0 and 0.31. This interval is rather large,
indicating that a significant amount of sampling is required to have credible
interaction networks. Fortunately, there are Bayesian methods which could be
used to complement available information and reduce this uncertainty
\citep{Cirtwill2018}.

%========================================================%
\section*{Conclusion}

Our representation of spatial variation of community structure offers a new
approach for the study of the biogeography of ecological networks. We see the
following key challenges and opportunities ahead in this exciting area of
research:

\textbf{1. New generation of network data}. Investigating spatial
variation of network structure will require high quality and highly replicated
network data. We have investigated one the most comprehensive spatial network
datasets we are aware of and nonetheless found immense gaps of knowledge in its
resolution. Species richness accumulates much faster than observations of
ecological interactions \citep{Poisot2012}. Each pair of species must be
observed several times in order to obtain reliable estimates of their interaction
probability.

\textbf{2. Estimation of the reliability of interactions}. We need
quantitative tools to estimate the confidence intervals around inferred
interaction probabilities, as well as estimators of the frequency of false
absences. Bayesian methods are promising to that end because we could use
information on the target species (e.g. if they are known as specialists or
generalists) to provide prior estimates of the interaction probability.

\textbf{3. From interaction probabilities to a distribution of network
properties}. Metrics are available to analyse the structure of probabilistic
networks \citep{Poisot2015c}. These metrics are useful as first approximation,
but they assume independence among interactions. This might not be the case in
nature because of the role of co-occurrence and shared environmental
requirements. We also need to better understand the distribution of network
properties arising from probabilistic interactions.

\textbf{4. Investigation of the environmental-dependence of ecological
interactions}. There is evidence that interactions can vary in space, but this
problem has not been investigated in a systematic fashion. The paucity of
currently available data precludes an extensive analysis of this question at
present.

\textbf{5. Effects of ecological interactions on co-occurrence}. We have
intentionally omitted the feedback of ecological interactions on co-occurrence
in this framework. As abundance can impact the occurrence of interactions, and,
conversely since interactions impact abundance \citep{Canard2014}, we could
reasonably expect that interactions will also influence co-occurrence. Theory
in this regard does exist for simple three-species modules
\citep{Cazelles2015}, but its extension to entire co-occurrence networks will
prove critical in the future, especially given the interest in using co-
occurrence to infer ecological interactions \citep{Morales-Castilla2015,
Morueta-Holme2016}.

%========================================================%
\section*{Acknowledgements}

This is a contribution to the working groups \emph {Continental-scale
variation of ecological networks} supported by the Canadian Institute for
Ecology and Evolution), and the \emph{Next Generation Data, Models, and
Theory Working Group}, supported by the Santa Fe Institute, the Betsy and
Jesse Fink Foundation, the ASU-SFI Center for Biosocial Complex Systems, and
NSF Grant PHY-1240192. DG also acknowledges financial support from NSERC-
Discovery grant program and Canada Research Chair program.

\newpage

%========================================================%
\bibliography{library}

\newpage
%========================================================%

%\begin{landscape}
\begin{table}[]
\centering

\caption{Example of data used to perform the analysis. Traditional species distribution models represent the occurrence of single species as a function of the environment, with models such as $X_i­~Temp$ performed one species after the other. Some also tried to represent both the abiotic and biotic niche, with models such as $X_i­~Temp + X_j$ (e.g. \citealt{GonzalezSalasar2013}), while the new joint species distribution models \citep{Ovaskainen2017} consider the entire community at once ($\mathbf{X}~Temp$) and aims at representing co-distribution ($X_{ij}$). Here the focus is on the occurrence of interactions, $L_{ij}$. The full model (see model 3, Tables 1 and 2) considers the result of two submodels, one representing the occurrence of interactions given co-occurrence, $L_{ij}~Temp$ (evaluated only on the subset of data where $X_{ij}=1$), and the other representing the co-occurrence, $X_{ij}~Temp$.}

\begin{tabular}{llllllll}
\hline

Location & $Sp_i$ & $Sp_j$ & $X_i$ & $X_j$ & $X_{ij}$ & $L_{ij}$ & Av. temp \\ \hline
Kühkopf 		& Phyllocolpa prussica  & Chrysocharis elongata & 1 & 1 & 1 & 1 &  8.25 \\
Kühkopf 		& Phyllocolpa leucapsis & Chrysocharis elongata & 0 & 1 & 0 & 0 &  8.25 \\
Ötztal 			& Phyllocolpa prussica  & Chrysocharis elongata & 0 & 0 & 0 & 0 & -1.50 \\
Ötztal 			& Phyllocolpa leucapsis & Chrysocharis elongata & 0 & 0 & 0 & 0 & -1.50 \\
Kilpisjärvi		& Phyllocolpa prussica  & Chrysocharis elongata & 0 & 1 & 0 & 0 & -1.92 \\
Kilpisjärvi		& Phyllocolpa leucapsis & Chrysocharis elongata & 0 & 1 & 0 & 0 & -1.92 \\
Wüstensachsen	& Phyllocolpa prussica  & Chrysocharis elongata & 1 & 1 & 1 & 0 &  5.58 \\
Wüstensachsen	& Phyllocolpa leucapsis & Chrysocharis elongata & 1 & 1 & 1 & 1 &  5.58 \\
Steinamoen      & Phyllocolpa prussica  & Chrysocharis elongata & 0 & 1 & 0 & 0 &  3.17 \\
Steinamoen      & Phyllocolpa leucapsis & Chrysocharis elongata & 0 & 1 & 0 & 0 &  3.17 \\

\hline
\end{tabular}
\end{table}
%\end{landscape}

\newpage
%========================================================%
\begin{landscape}
\begin{table}[]
\centering
\caption{Summary of model comparison for the interaction between the leaf
folder \textit{Phyllocolpa prussica}) and the parasitoid \textit{Chrysocharis
elongata}}. Note that $E$ is a short notation including both temperature and precipitation.
\begin{tabular}{llllll}
\hline
	\# & Metaweb model 						& Co-occurrence model 			& LL 	& npars & AIC \\ \hline
	1 & $P(L_{ijy})$ 						& $P(X_{iy},X_{jy}\lvert E_y)$ 		& -71.1 & 6 	& 154.2 \\
	2 & $P(L_{ijy} \lvert  X_{iy}, X_{jy})$ 		& $P(X_{iy},X_{jy}\lvert E_y)$ 		& -65.7 & 6 	& 143.4 \\
	3 & $P(L_{ijy} \lvert  X_{iy}, X_{jy}, E_y)$ & $P(X_{iy},X_{jy}\lvert E_y)$ 		& -65.6 & 10 	& 151.3 \\ \hline
	4 & $P(L_{ijy} \lvert  X_{iy}, X_{jy}, E_y)$ & $P(X_{iy},X_{jy})$ 			& -84.5 & 6 	& 183 \\
	5 & $P(L_{ijy} \lvert  X_{iy}, X_{jy}, E_y)$ & $P(X_{iy})P(X_{jy})$ 			& -80.7 & 7 	& 173.4 \\
	6 & $P(L_{ijy} \lvert  X_{iy}, X_{jy}, E_y)$ & $P(X_{iy}\lvert E_y)P(X_{jy}\lvert E_y)$ 	& -68.8 & 15 	& 167.6 \\
\hline
\end{tabular}
\end{table}
\end{landscape}

\newpage
%========================================================%

\begin{landscape}
\begin{table}[]
\centering
\caption{Summary of model comparison for the interaction across all pairs of \textit{Salix}, gallers and parasitoids.}
\begin{tabular}{lllllll}
\hline
	Interaction & \# & Metaweb model & Co-occurrence model & LL & npars & AIC \\ \hline
	\textit{Salix}-Galler & 1 & $P(L_{ijy})$ & $P(X_{iy},X_{jy}\lvert E_y)$ & -6022.1 & 7548 & 27140.3 \\
	 & 2 & $P(L_{ijy}\lvert X_{iy},X_{jy})$ & $P(X_{iy},X_{jy}\lvert E_y)$ & -5547.9 & 7548 & 26191.8 \\
	 & 3 & $P(L_{ijy} \lvert  X_{iy}, X_{jy}, E_y)$ & $P(X_{iy},X_{jy}\lvert E_y)$ & -5364.0 & 12580 & 35888.0 \\
	 & 4 & $P(L_{ijy} \lvert  X_{iy}, X_{jy}, E_y)$ & $P(X_{iy},X_{jy})$ & -5998.4 & 8806 & 30287.2 \\
	 & 5 & $P(L_{ijy} \lvert  X_{iy}, X_{jy}, E_y)$ & $P(X_{iy})P(X_{jy})$ & -6636.1 & 7548 & 27092.7 \\
	 & 6 & $P(L_{ijy} \lvert  X_{iy}, X_{jy}, E_y)$ & $P(X_{iy}\lvert E_y)P(X_{jy}\lvert E_y)$ & -6002.9 & 18870 & 49745.7 \\ \hline
	Galler-Parasitoid & 1 & $P(L_{ijy})$ & $P(X_{iy},X_{jy}\lvert E_y)$ & -22068.1 & 19206 & 82548.2 \\
	 & 2 & $P(L_{ijy} \lvert  X_{iy}, X_{jy})$ & $P(X_{iy},X_{jy}\lvert E_y)$ & -19504.8 & 19206 & 77421.6 \\
	 & 3 & $P(L_{ijy} \lvert  X_{iy}, X_{jy}, E_y)$ & $P(X_{iy},X_{jy}\lvert E_y)$ & -20217.0 & 32010 & 104454.1 \\
	 & 4 & $P(L_{ijy} \lvert  X_{iy}, X_{jy}, E_y)$ & $P(X_{iy},X_{jy})$ & -19591.3 & 22407 & 77594.5 \\
	 & 5 & $P(L_{ijy} \lvert  X_{iy}, X_{jy}, E_y)$ & $P(X_{iy})P(X_{jy})$ & -22491.5 & 19206 & 89796.9 \\
	 & 6 & $P(L_{ijy} \lvert  X_{iy}, X_{jy}, E_y)$ & $P(X_{iy}\lvert E_y)P(X_{jy}\lvert E_y)$ & -18936.9 & 48015 & 133903.7 \\
\hline
\end{tabular}
\end{table}
\end{landscape}

\newpage
%========================================================%
\section*{Figure legends}

%------------------------
\subsection*{Figure 1}

\textbf{Non-random sampling of the metaweb}. Network assembly can be viewed
as a sampling process of the regional pool of potential interactions. Species
(indicated by colored nodes) are sampled first, and among the species found in
the local network, only some interactions (indicated by blue links) occur.
We characterize these sampling processes with the quantitative framework
proposed in this paper. As a concrete illustration of metaweb sampling, we
here show a local interaction network among \textit{Salix} (bottom/green), gallers
(center/blue), and parasitoids (top/red). The metaweb was constructed by
aggregating interactions observed across 370 local networks. Nodes were distributed
to minimize crossing of interactions.

%------------------------
\subsection*{Figure 2}

\textbf{Visual representation of the integrated niche}. In biogeography, the
niche is considered the set of environmental conditions where the intrinsic
growth rate $r$ is positive \citep{Holt2009a}. The horizontal axis represents
an environmental gradient impacting the growth of a focal species (red point).
The location of each species (grey points) along this gradient represents their
optimum, and the vertical dotted lines represent the limits of the Grinnellian
niche of the focal species. In food web ecology, the Eltonian niche represents
the location of a species in the food web, as determined by its niche position
($n$) and its niche optimum ($c$). The vertical axis represents a niche
gradient, for example a trait such as body size. The location of each species
along this gradient represents their niche position. The focal species will
feed only on host/prey species occupying niche locations within a given interval
around the optimum, represented by the horizontal lines. The integrated
Grinnellian and Eltonian niche corresponds to the square in the middle where
an interaction is possible owing to a match of traits and spatial
distribution. The central square represents the area where the joint
probability of observing co-occurrence and interactions is positive.

%------------------------
\subsection*{Figure 3}

\textbf{Probabilistic representation of the interaction probability between a
leaf folder (\textit{Phyllocolpa prussica}) and a parasitoid
(\textit{Chrysocharis elongata}) across gradients of annual average
temperature and annual precipitation}. The colour gradient represents the
probability of observing the event, from 0 (white) to 1 (black). The
representation is based on predictions from Model 3 (see Table 1). In the left
panel, open circles represent the absence of both species, whereas closed
circles represent co- occurrence and plus signs the occurrence of only one of
the two species. In the other two panels, open circles represent co-occurrence
but an absence of interaction and closed circles the occurrence of an
interaction.

%------------------------
\subsection*{Figure 4}

\textbf{Probabilistic representation of the interaction probability between a
leaf folder (\textit{Phyllocolpa prussica}) and a parasitoid
(\textit{Chrysocharis elongata}) across Europe}. Climate data are extracted
from WorldClim over a 1km x 1km grid and predictions made according to the
model illustrated at Fig. 3. Colour gradient ranges from low values (dark
blue) to high values (yellow).

% %------------------------
% \subsection*{Figure 5}

% \textbf{Representation of the \textit{Salix}-galler and galler-parasitoid metawebs}.
% Black cells indicate species pairs for which at least one interaction was
% recorded, white cells indicate absence of recorded interactions and grey
% cells show pairs of species never detected at the same site (and hence species
% pairs for which we have no information on whether they would interact should
% they co-occur).

%------------------------
\subsection*{Figure 5}

\textbf{Mapping the distribution of species richness, the number of links and
connectance across Europe}. The representation is based on predictions from
Model 3 (see Table 2) using climate data extracted from WorldClim over a 1km x
1km grid. Species richness is obtained by summation of individual occurrence
probabilities, and link density by summation of interaction probabilities.
Color gradient ranges from low values (dark blue) to high values (yellow).

\newpage

%========================================================%
%------------------------
\subsection*{Figure 1}

\begin{figure}[ht!]
\centering\includegraphics[width=1\textwidth]{figures/mw_sampling.png}
\end{figure}

\newpage

%------------------------
\subsection*{Figure 2}

\begin{figure}[ht!]
\centering\includegraphics[width=0.6\textwidth]{figures/integrated_niche.png}
\end{figure}

\newpage

%------------------------
\subsection*{Figure 3}

\begin{figure}[ht!]
\centering\includegraphics[width=0.8\textwidth]{figures/example_pair.png}
\end{figure}

\newpage

%------------------------
\subsection*{Figure 4}

\begin{figure}[ht!]
\centering\includegraphics[width=0.8\textwidth]{figures/map_pair.png}
\end{figure}

\newpage

% %------------------------
% \subsection*{Figure 5}

% \begin{figure}[ht!]
% \centering\includegraphics[width=0.8\textwidth]{figures/mw_holes}
% \end{figure}

% \newpage

%------------------------
\subsection*{Figure 5}

\begin{figure}[ht!]
\centering\includegraphics[width=0.8\textwidth]{figures/map_connectance.png}
\end{figure}

\newpage

%========================================================%
\end{document}